---
title: "Week 02 - dplyr exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 01: select
Use [diamonds](https://ggplot2.tidyverse.org/reference/diamonds.html) example table from `ggplot2` package.

Select only `price`, `cut`, and `color` columns. Use piping and `select`.
```{r exercise 01-1}
library(tidyverse)
data <- diamonds

diamonds %>% 
  select(price, cut, color)

```

Select only `price`, `cut`, and `color` columns. Use base R (no piping or select).
```{r exercise 01-2}
diamonds$cut 
diamonds$price
diamonds$color

ccp <- c(diamonds$cut, diamonds$price,diamonds$color)

cpp <- data.frame(ccp)

diamonds[,c(2,3,7)]
```

Select all columns except for `x`, `y`, and `z`. Use negative selection and piping.
```{r exercise 01-3}
diamonds %>% 
  select(-x,-y,-z)
```

Check out the help page for the `select()` function. Use `:` notation to select columns from `carat` till `clarity`.
```{r exercise 01-4}
help("select")

diamonds %>% 
  select(,1:4)
```

And now select these same columns by capitalising on the fact that all their names (and none of the others) start with the letter "c".

```{r exercise 01-5}

diamonds %>% 
  select(starts_with('c'))

```

BONUS: How would you select these 4 columns in base R, without simply listing their names?

```{r exercise 01-6}
diamonds [,1:5]
```


## Exercise 02: conditions 1
Examine the code and decide on whether a condition will be `TRUE` or `FALSE`. Write down the answer and run the chunk to check the answer.

**Answer 1:**
```{r exercise 02-1}
x <- 4
(x > 3)
#TRUE
```

**Answer 2:**
```{r exercise 02-2}
x <- 4
y <- 5
(x != y)
#TRUE
```
## Exercise 03: conditions 2
Examine the code and decide on whether a condition will be `TRUE` or `FALSE`. Write down the answer and run the chunk to check the answer.

**Answer 1:**
```{r exercise 03-1}
x <- 4
!(x < 3)
#TRUE
```

**Answer 2:**
```{r exercise 03-2}
x <- 4
y <- 5
!(x != y)
#FALSE
```

## Exercise 04: conditions 3
Examine the code and decide on whether a condition will be `TRUE` or `FALSE`. Write down the answer and run the chunk to check the answer.

**Answer 1:**
```{r exercise 03-1}
x <- 4
(x < 3) | (x >=4)
#TRUE
```

**Answer 2:**
```{r exercise 04-2}
x <- 4
y <- 6
(x == y) | ((x < 5) & (y > 5))
#TRUE
```

**Answer 3:**
```{r exercise 04-3}
x <- 3
y <- 3
(x != y) & ((x <= 3) & (y >= 3))
#FALSE
```

**Answer 4:**
```{r exercise 04-4}
x <- 3
y <- 3
(x != y) | ((x <= 3) & (y >= 3))
#TRUE
```
## Exercise 05: conditions
Examine the code and decide on whether a condition will be `TRUE` or `FALSE`. Write down the answer and run the chunk to check the answer.

**Answer 1:**
```{r exercise 05-1}
x <- seq(-4, 4, 2)
(x == 0)

#FALSE FALSE  TRUE FALSE FALSE

help("seq")
```
**Answer 2:**
```{r exercise 05-2}
x <- seq(-4, 4, 2)
y <- 1:5
(x + y <= 0)

#TRUE  TRUE FALSE FALSE FALSE

```

**Answer 3:**
```{r exercise 05-3}
x <- seq(-4, 12, 4)
y <- 1:5
(x > y)

#FALSE FALSE  TRUE  TRUE  TRUE
```
**Answer 4:**
```{r exercise 05-4}
x <- seq(-4, 12, 4)
y <- 1:5
(x > y) & ((x < 10) | (y < 3))

#FALSE FALSE  TRUE  TRUE FALSE

```
## Exercise 06: logical indexing

Use logical indexing to values of `x` that are above 3.
```{r exercise 06-1}
x <- -3:7
# Your code goes here
x[x>3]

```

Use logical indexing to extract values of `x` when it is higher than value in `y`
```{r exercise 06-2}
x <- -3:7
y <- seq(5, -2, length.out = 11)
x[x>y]

# Your code goes here
```

Use logical indexing to extract values of `x` that are greater than the mean value of x (avoid using intermediate variables).
```{r exercise 06-3}
x <- -3:7
mean_x <- mean(x)


# Your code goes here
x[x>(mean(x))]
```

## Exercise 07: filter
Use [msleep](https://ggplot2.tidyverse.org/reference/msleep.html) example table from `ggplot2` package. 

Filter the table to keep only `"domesticated"` animals (variable name is `conservation`).
```{r exercise 07-1}
data <- msleep %>% 
  filter(conservation == 'domesticated')
```

Filter the table to keep only the omnivores (`vore`) that sleep more than 10 hours in total.
```{r exercise 07-2}
msleep %>% 
  filter(vore == 'omni') %>% 
  filter(sleep_total>10)
```

Filter the table to exclude animals with body weight above `50`.
```{r exercise 07-3}
msleep %>% 
  filter(bodywt<=50)

```

Filter the table to exclude animals with body weight below [median](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/median.html) body weight. Hint, you do not need a temporary variable. 
```{r exercise 07-4}

msleep %>% 
  filter(bodywt < median(bodywt))
```


## Exercise 08: arrange
Use [presidential](https://ggplot2.tidyverse.org/reference/presidential.html) example table from _ggplot2_ package.

Arrange the table by party.
```{r exercise 08-1}

```

Arrange the table by start date in the descending order.
```{r exercise 08-2}

```

Arrange the table by party (descending order) and president's name (ascending order).
```{r exercise 08-3}

```


## Exercise 9: mutate

Use [msleep](https://ggplot2.tidyverse.org/reference/msleep.html) example table from `ggplot2` package.

Compute new variable / column with sleep to body weight ratio.
```{r exercise 10-1}
```

Compute new variable / column `is_big` of logical value that are `TRUE` if body weight is above `50`.
```{r exercise 10-2}
```

Compute three new variables / columns. First, compute the _mean_ and _standard deviation_ of total sleep variable. You will get the same value in all the rows, make sure that your columns are _not_ called `mean` and `sd`. Then compute a z-score of the total sleep: $Z(x) = (x - mean(x)) / sd(x)$. Put all computations into a single mutate code.
```{r exercise 10-3}
```

Compute z-score for body weight variable, do not store mean and standard deviation in extra columns but use them directly.
```{r exercise 10-4}
```

