---
title: "datasauRus"
author: "Kasia Banas"
date: "2024-09-26"
output: html_document
---

Install and load the datasauRus library:

```{r}
# install.packages("datasauRus")
library(datasauRus)
```

Load the tidyverse: 
```{r}
library(tidyverse)
```

`datasaurus_dozen` is a dataset demonstrating the utility of visualization. 
These 12 datasets are equal in standard measures: mean, standard deviation, 
and Pearson's correlation. But they look very different
Run the chunk below to get more info about this dataset:

```{r}
?datasaurus_dozen
```

See what datasets are in the datasaurus_dozen:

```{r}
datasaurus_dozen %>% 
  distinct(dataset)
```

## Exercise 1
Pick 2 datasets and plot them next to one another
(I'd recommend scatterplot, use facets if you like)

view('star')
view('dino')

view(datasaurus_dozen)

datasaurus_data = datasaurus_dozen

datasaurus_data %>% 
  ggplot(aes(x=x, y=y, group = 'away')) +
  geom_point(colour = 'darkgreen') +
  facet_wrap(~dataset, scales = "free")
  
datasaurus_data %>% 
  ggplot(aes(x=x, y=y, colour = dataset)) +
  geom_point() +
  facet_wrap(~dataset, scales = "free")
  
dino_data = datasaurus_data %>%
  filter(dataset == 'dino')
  
star_data = datasaurus_data %>%
  filter(dataset == 'star')
  
dino_data %>% 
ggplot(aes(x=x, y=y)) +
geom_point()
  
star_data %>% 
ggplot(aes(x=x, y=y)) +
geom_point()

both_data = datasaurus_data %>% 
  filter(dataset == "dino" | dataset == "star")
  
both_data %>% 
  ggplot(aes(x=x, y=y, colour = dataset)) +
  geom_jitter() +
  facet_wrap(~dataset)
  
datasaurus_data %>% 
  filter(dataset == "dino" | dataset == "star") %>%
  ggplot(aes(x=x, y=y, colour = dataset)) +
  geom_jitter() +
  facet_wrap(~dataset)
  

## Exercise 2
For these same two datasets, find the mean and standard deviation
for the x and y variable in each. What did you find? 

sd(both_data$x)
sd(both_data$y)

mean(both_data$x)
mean(both_data$y)

datasaurus_data %>% 
  filter(dataset == "dino" | dataset == "star") %>% 
  group_by(dataset) %>% 
  summarise(mean_x = mean(x), sd_x =(sd(x)), mean_y = mean(y), sd_x = sd(y))

